\tit

### 2021-02-22 paper.tex
Use ◇ for diamond

%% I prefer ∅ in XITSMath-Regular, but I don't know how to make setmathfont
%% work with acmart.
\ifacm
\newunicodechar{◬}{\raisebox{0.02ex}{\ensuremath{_{\diamond}\hspace{-0.323em}}}}
\newunicodechar{▣}{\raisebox{-0.12ex}{\ensuremath{\scaleobj{0.4}{\Box}}\hspace{-0.238em}}}
\else
\setmathfont{XITSMath-Regular.otf}
\newunicodechar{◬}{\raisebox{0.23ex}{\ensuremath{_{\diamond}\hspace{-0.358em}}}}
\newunicodechar{▣}{\raisebox{-0.13ex}{\ensuremath{\scaleobj{0.3}{\Box}}\hspace{-0.23em}}}
\newunicodechar{∷}{\ensuremath{\mathbin{\hspace{-0.2em}:\hspace{-0.27em}:\hspace{-0.15em}}}}
\fi

@@

The ``symbolic differentiation'' style reflects language-building vocabulary (\figref{Lang-ops}) into an inductive data type (of modestly extended regular expressions), while \AF{ν} and \AF{δ} become functions defined by pattern-matching on that data type.
For convenience, we can use the same names as in \secref{Specifying Languages} for these new counterparts, while referring to the original versions via the module prefix ``{◬.}\hspace{0.05em}''.
The result is shown in \figrefdef{symbolic-api}{Regular expressions (inductive)}{\api} along with a function {\AF{⟦\_⟧}} that converts a symbolic language representation to a computable parser.

@@

\secref{Symbolic Differentiation} embodies one choice of reflecting functions into a data representation.
Now consider the dual strategy of ``automatic differentiation''.
This time, reflect \AF{ν} and \AF{δ} into a \emph{coinductive} data type of tries\footnote{The classic trie (``prefix tree'') data structure was introduced by \citet{Thue1912Gegenseitige} to represent sets of strings \citep[Section 6.3]{Knuth1998ACP3}, later generalized to arbitrary non-nested algebraic data types \citep{Connelly1995GenTrie}, and then from sets to functions \citep{Hinze2000GGT}.}, while redefining the language-building vocabulary as functions on that data type defined by \emph{copatterns} \citep{AbelPientka2016}.
Again, we will use the same names as in \secref{Specifying Languages}, referring to the original versions via the module prefix ``{◬.}\hspace{0.05em}''.
The result is shown in \figrefdef{automatic-api}{Tries (coinductive)}{\api}.

@@

Fortunately, there is a simple way to encapsulate indexed types into non-indexed types, namely existential quantification.
For instance, {\AF{∃} \AD{Dec}} and {\AF{∃} \AD{(\ARe{Lang} \APo ∞)}} are non-indexed types.
For a type family {\AB F \AK : \AB A \AK → \APT{Set} \AB ℓ}, values of type {\AF ∃ \AB F} are pairs {(\AB a , \AB b)} with {\AB a \AK : \AB A} and {\AB b \AK : \AB F \AB a}.
We can also encapsulate dependently typed \emph{operations} on indexed types into simply typed operations on the (simply typed) existentially wrapped versions of those types.
For instance, consider language union from \figref{symbolic-api}:
\begin{code}
% Copied from latex/Symbolic.tex
\>[2]\AgdaOperator{\AgdaInductiveConstructor{\AgdaUnderscore{}∪\AgdaUnderscore{}}}%
\>[7]\AgdaSymbol{:}\AgdaSpace{}%
\AgdaDatatype{Lang}%
\>[15]\AgdaGeneralizable{P}%
\>[18]\AgdaSymbol{→}\AgdaSpace{}%
\AgdaDatatype{Lang}\AgdaSpace{}%
\AgdaGeneralizable{Q}%
\>[28]\AgdaSymbol{→}\AgdaSpace{}%
\AgdaDatatype{Lang}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaGeneralizable{P}%
\>[39]\AgdaOperator{\AgdaFunction{◬.∪}}%
\>[44]\AgdaGeneralizable{Q}\AgdaSymbol{)}\<%
\end{code}
Referring to the type and operations of \figreftwo{Lang-ops}{symbolic-api} via the module prefixes ``{◬.}\hspace{0.05em}'' and ``{▣.}\hspace{0.05em}'' respectively, define \agda{wrapped-Lang}
Then wrap union as follows:
\agda{wrapped-union}
The type-level language components of the arguments and result are given explicitly here but could instead be inferred by the compiler.
\rnc\source{Existential}
Indexed operations of all arities for a type family \AB{F} can be systematically wrapped in this style, as shown in
\figrefdef{inj}{Existential wrappers}{\mathindent0ex\agda{inj}}, again with automatically inferable components given explicitly for clarity.
Then wrapped union can be defined simply as {\AF{\_∪\_} \AS = \AF{inj₂} \AIC{▣.\_∪\_}} and likewise for the other operations.

>>>>

%% I prefer ∅ in XITSMath-Regular, but I don't know how to make setmathfont
%% work with acmart.
\ifacm
\newunicodechar{◇}{\raisebox{0.02ex}{\ensuremath{_{\diamond}\hspace{-0.323em}}}}
\newunicodechar{▣}{\raisebox{-0.12ex}{\ensuremath{\scaleobj{0.4}{\Box}}\hspace{-0.238em}}}
\else
\setmathfont{XITSMath-Regular.otf}
\newunicodechar{◇}{\raisebox{0.23ex}{\ensuremath{_{\diamond}\hspace{-0.358em}}}}
\newunicodechar{▣}{\raisebox{-0.13ex}{\ensuremath{\scaleobj{0.3}{\Box}}\hspace{-0.23em}}}
\newunicodechar{∷}{\ensuremath{\mathbin{\hspace{-0.2em}:\hspace{-0.27em}:\hspace{-0.15em}}}}
\fi

@@

The ``symbolic differentiation'' style reflects language-building vocabulary (\figref{Lang-ops}) into an inductive data type (of modestly extended regular expressions), while \AF{ν} and \AF{δ} become functions defined by pattern-matching on that data type.
For convenience, we can use the same names as in \secref{Specifying Languages} for these new counterparts, while referring to the original versions via the module prefix ``{◇.}\hspace{0.05em}''.
The result is shown in \figrefdef{symbolic-api}{Regular expressions (inductive)}{\api} along with a function {\AF{⟦\_⟧}} that converts a symbolic language representation to a computable parser.

@@

\secref{Symbolic Differentiation} embodies one choice of reflecting functions into a data representation.
Now consider the dual strategy of ``automatic differentiation''.
This time, reflect \AF{ν} and \AF{δ} into a \emph{coinductive} data type of tries\footnote{The classic trie (``prefix tree'') data structure was introduced by \citet{Thue1912Gegenseitige} to represent sets of strings \citep[Section 6.3]{Knuth1998ACP3}, later generalized to arbitrary non-nested algebraic data types \citep{Connelly1995GenTrie}, and then from sets to functions \citep{Hinze2000GGT}.}, while redefining the language-building vocabulary as functions on that data type defined by \emph{copatterns} \citep{AbelPientka2016}.
Again, we will use the same names as in \secref{Specifying Languages}, referring to the original versions via the module prefix ``{◇.}\hspace{0.05em}''.
The result is shown in \figrefdef{automatic-api}{Tries (coinductive)}{\api}.

@@

Fortunately, there is a simple way to encapsulate indexed types into non-indexed types, namely existential quantification.
For instance, {\AF{∃} \AD{Dec}} and {\AF{∃} \AD{(\ARe{Lang} \APo ∞)}} are non-indexed types.
For a type family {\AB F \AK : \AB A \AK → \APT{Set} \AB ℓ}, values of type {\AF ∃ \AB F} are pairs {(\AB a , \AB b)} with {\AB a \AK : \AB A} and {\AB b \AK : \AB F \AB a}.
We can also encapsulate dependently typed \emph{operations} on indexed types into simply typed operations on the (simply typed) existentially wrapped versions of those types.
For instance, consider language union from \figref{symbolic-api}:
\begin{code}
% Copied from latex/Symbolic.tex
\>[2]\AgdaOperator{\AgdaInductiveConstructor{\AgdaUnderscore{}∪\AgdaUnderscore{}}}%
\>[7]\AgdaSymbol{:}\AgdaSpace{}%
\AgdaDatatype{Lang}%
\>[15]\AgdaGeneralizable{P}%
\>[18]\AgdaSymbol{→}\AgdaSpace{}%
\AgdaDatatype{Lang}\AgdaSpace{}%
\AgdaGeneralizable{Q}%
\>[28]\AgdaSymbol{→}\AgdaSpace{}%
\AgdaDatatype{Lang}\AgdaSpace{}%
\AgdaSymbol{(}\AgdaGeneralizable{P}%
\>[39]\AgdaOperator{\AgdaFunction{◇.∪}}%
\>[44]\AgdaGeneralizable{Q}\AgdaSymbol{)}\<%
\end{code}
Referring to the type and operations of \figreftwo{Lang-ops}{symbolic-api} via the module prefixes ``{◇.}\hspace{0.05em}'' and ``{▣.}\hspace{0.05em}'' respectively, define \agda{wrapped-Lang}
Then wrap union as follows:
\agda{wrapped-union}
The type-level language components of the arguments and result are given explicitly here but could instead be inferred by the compiler.
\rnc\source{Existential}
Indexed operations of all arities for a type family \AB{F} can be systematically wrapped in this style, as shown in
\figrefdef{inj}{Existential wrappers}{\mathindent0ex\agda{inj}}, again with automatically inferable components given explicitly for clarity.
Then wrapped union can be defined simply as {\AF{\_∪\_} \AS = \AF{inj₂} \AIC{▣.\_∪\_}} and likewise for the other operations.
